SMT-Based Generation of Symbolic Automata \thanks{This work was partially funded by the Associated Team FM4CPS between INRIA and ECNU, Shanghai}

### 2019-11-05 2017-Xudong-SMT/ActaInformatica.tex

\subsection{Running example: the Enable operator}
\ERIC{To illustrate our definitions and results, we will use a small running example, coming from the specification language LOTOS \cite{LotosISO89}.
  Suppose we have two unspecified processes $P$ and $Q$. In the Enable expression ``$P>>Q$'', an \texttt{exit} statement within $P$ terminates the current process and pass the control to $Q$, carrying a value $x$ that is captured by the \texttt{accept} statement of $Q$.
  In the next section we will define pNet encodings of the Enable operator. Let us start by defining the algebra presentation that will be used in these pNets.
}

@@

\ERIC{The $EnableData$ pNet will use the same algebra presentation.}
%\QIN{
%Another possible encoding of such ``Enable'' operator, called $EnableData$, uses a data-oriented style, with a single state controller $CData$, and a state-variable $v$, with values in $\{0,1\}$. As there is no more constructor required, we can use the same algebra presentation as $EnableState$ in this case.
%}
\end{example}

@@

\begin{example}[PLTS of the Controller in the ``Enable'' Operator]
\ERIC{
The PLTS of the controller $CState$ from Example \ref{example:EnableOperator} is  $\langle S,s_0, \to\rangle$, where:
\begin{itemize}
	\item[$\bullet$] $S=\{0, 1\}$, $s_0=0$;
	\item[$\bullet$] $\to=\{0\xrightarrow{l}0, 0\xrightarrow{\delta}1, 1\xrightarrow{r}1\}$.
\end{itemize}
Featuring guards and assignments in the label of transitions, the PLTS of the controller $CData$ is $\langle S,s_0, \to\rangle$ where:
\begin{itemize}
	\item[$\bullet$] $S=\{0, 1\}$, $s_0=0$, $vars(s_0)=\{v\}$, $\texttt{init}\ \{v:=0\}$;
	\item[$\bullet$] $\to=\{0\xrightarrow{l, [v=0]}0, 0\xrightarrow{\delta,  [v=0], \{v\leftarrow1\}}1, 1\xrightarrow{r, [v=1]}1\}$.
\end{itemize}
}
\ERIC{The PLTSs contained in $EnableStateCompLeft$ are similar to the PLTS $CState$, but their action constants have been renamed, ensuring disjoint state variable sets.}
\end{example}

@@

\ERIC{Each call to the SMT engine will return either SAT or UNSAT results, or may fail due to time bounds. Also, the accuracy of a 'SAT' result may depend on the characteristics of the specific theory used to axiomatise the data domains and operators. Such theories may be intrinsically undecidable, or may not properly handled by the SMT engine (e.g. for recursive datatypes).
However, while building the open automaton, that is a symbolic representation of the pNet behavior, it is not logically incorrect to keep potentialy unsatisfiable transitions, so we will treat undecidable results, and time bounds failures, as SAT. So the open automaton, may have more transitions and states that the theoretical minimum.}

>>>>

\subsection{Running example: the Enable operator}
\ERIC{To illustrate our definitions and results, we will use a small running example, coming from the specification language LOTOS \cite{LotosISO89}.
  Suppose we have two unspecified processes $P$ and $Q$. In the Enable expression ``$P>>Q$'', an \texttt{exit} statement within $P$ terminates the current process and passes the control to $Q$, carrying a value $x$ that is captured by the \texttt{accept} statement of $Q$.
  In the next section we will define pNet encodings of the Enable operator. Let us start by defining the algebra presentation that will be used in these pNets.
}

@@

\ERIC{The $EnableData$ pNet will use the same algebra presentation.}\QIN{The main difference between $EnableState$ and $EnableData$ lies in the construction of their controller $CState$ and $CData$ which we will discuss in the next section.}
%\QIN{
%Another possible encoding of such ``Enable'' operator, called $EnableData$, uses a data-oriented style, with a single state controller $CData$, and a state-variable $v$, with values in $\{0,1\}$. As there is no more constructor required, we can use the same algebra presentation as $EnableState$ in this case.
%}
\end{example}

@@

\begin{example}[PLTS of the Controller in the ``Enable'' Operator]
\ERIC{
The PLTS of the controller $CState$ from Example \ref{example:EnableOperator} is  $\langle S,s_0, \to\rangle$, where:
\begin{itemize}
	\item[$\bullet$] $S=\{0, 1\}$, $s_0=0$;
	\item[$\bullet$] $\to=\{0\xrightarrow{l}0, 0\xrightarrow{\delta}1, 1\xrightarrow{r}1\}$.
\end{itemize}
}
\QIN{
Featuring guards and assignments in the label of transitions, the PLTS of the controller $CData$ is $\langle S,s_0, \to\rangle$ where:
\begin{itemize}
	\item[$\bullet$] $S=\{0\}$, $s_0=0$, $vars(s_0)=\{v\}$, $\texttt{init}\ \{v:=0\}$;
	\item[$\bullet$] $\to=\{0\xrightarrow{l, [v=0]}0, 0\xrightarrow{\delta,  [v=0], \{v\leftarrow1\}}0, 0\xrightarrow{r, [v=1]}0\}$.
\end{itemize}
}
\ERIC{The PLTSs contained in $EnableStateCompLeft$ are similar to the PLTS $CState$, but their action constants have been renamed, ensuring disjoint state variable sets.}
\end{example}

@@

\ERIC{Each call to the SMT engine will return either SAT or UNSAT results, or may fail due to time bounds. Also, the accuracy of a `SAT' result may depend on the characteristics of the specific theory used to axiomatise the data domains and operators. Such theories may be intrinsically undecidable, or may not properly handled by the SMT engine (e.g. for recursive datatypes).
However, while building the open automaton, that is a symbolic representation of the pNet behavior, it is not logically incorrect to keep potentially unsatisfiable transitions, so we will treat undecidable results, and time bounds failures, as SAT. So the open automaton, may have more transitions and states that the theoretical minimum.}
